---
title: "Practical Machine Learning Assignment"
author: "Prabeeti Bulani"
date: "10/11/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 1. Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

# 2. Objective
In this project, my objective is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants and predict the manner in which these particiapnts have prerformed these exercises. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
#Data Source & References:- #http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har 
#(refer-Weight Lifting Exercise Data)
#Training data - https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
#Test Set-https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
 
 
 # 3. Data Loading and Exploratory Analysis
```{r}
# Following library are used
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
library(janitor)
set.seed(123)
# Data Extraction
Training_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
Testing_url  <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
Training_data<-"pml-traininig.csv"
Testing_data<-"pml-testing.csv"

if(!file.exists(Training_data))
{
    download.file(Training_url,destfile = Training_data,method="curl")
}
training <- read.csv(Training_data, header = TRUE, sep = ",", na.strings=c("NA","#DIV/0!",""))
dim(training)
if(!file.exists(Testing_data))
{
    download.file(Testing_url,destfile = Testing_data,method="curl")
}
testing  <- read.csv(Testing_data, header = TRUE, sep = ",", na.strings=c("NA","#DIV/0!",""))
dim(testing)
# create a partition using caret with the training dataset on 60,40 ratio
inTraining  <- createDataPartition(training$classe, p=0.6, list=FALSE)

Training_dataset <- training[inTraining, ]

Test_dataset  <- training[-inTraining, ]
# Exploring the data in training and test data
dim(Training_dataset)
dim(Test_dataset)
# Training and testing dataset both have 160 variables, with records count divided as 60:40 ratio
str(Training_dataset)

# Data cleansing - cleaning the data for NA, The Near Zero variance (NZV) and ID variables 
Training_nearZeroVar <- nearZeroVar(Training_dataset)
Training_dataset <- Training_dataset[, -Training_nearZeroVar]
Test_nearZeroVar <- nearZeroVar(Test_dataset)
Test_dataset  <- Test_dataset[, -Test_nearZeroVar]
label <- apply(Training_dataset, 2, function(x) mean(is.na(x))) > 0.95
Training_dataset <- Training_dataset[, -which(label, label == FALSE)]
label_test <- apply(Test_dataset, 2, function(x) mean(is.na(x))) > 0.95
Test_dataset <- Test_dataset[, -which(label_test, label_test == FALSE)]


Training_dataset <- Training_dataset[ , -(1:5)]
Test_dataset <- Test_dataset[ , -(1:5)]

dim(Training_dataset)
dim(Test_dataset)

# Training and testing dataset  have 54 and 54  variables respectively after cleansing.
```


# 4.Predictive Modelling Processes

# 4.1 Decision Tree Model
```{r}

# DEcision tree build
set.seed(123)
model_decisionTree <- rpart(classe ~ ., Training_dataset, method="class")
fancyRpartPlot(model_decisionTree)

# Prediction with decision tree model
predict_DT<- predict(model_decisionTree, Test_dataset, type = "class")
confusionMatrix_DT <-confusionMatrix(predict_DT, Test_dataset$classe)
confusionMatrix_DT
# plot confusion matrix results
plot(confusionMatrix_DT$table, col = confusionMatrix_DT$byClass, 
     main = paste("Decision Tree - Accuracy =",
                  round(confusionMatrix_DT$overall['Accuracy'], 4)))
```

# 4.2 Randon Forest Model

```{r}

# Random Forest build
set.seed(123)
train_ctrl <-  trainControl(method = "cv", classProbs=TRUE,savePredictions=TRUE,allowParallel=TRUE, number = 10)
model_randomForest <- randomForest(classe ~ ., data = Training_dataset, method = "rf", importance = T, trControl = train_ctrl,na.action = na.exclude)


# Prediction with random forest model
predict_RF<- predict(model_randomForest, Test_dataset, type = "class")
confusionMatrix_RF <-confusionMatrix(predict_RF, Test_dataset$classe)
confusionMatrix_RF
# plot confusion matrix results for random forest
plot(model_randomForest)

```

# 4.3 Generalised Boosting Model

```{r}

# Generalised Boosting Model  build
set.seed(123)
gbm_control <- trainControl(method = "repeatedcv", number = 5, repeats =1,verboseIter = FALSE)
model_Boosting <- train(classe ~ ., method = "gbm",data= Training_dataset,
                    verbose = FALSE,
                    trControl = gbm_control)

model_Boosting

plot(model_Boosting)
# Prediction with Generalised Boosting Model model
predict_GBM<- predict(model_Boosting, Test_dataset)
confusionMatrix_GBM <-confusionMatrix(predict_GBM, Test_dataset$classe)
confusionMatrix_GBM

```
# Conclusion
#Accuracy obtained from 3 models are as below:-
#1. Decision Tree Model -0.8278
#2. Random Forest Model - 0.9963
#3. Boosting Model - 0.9862

Out of these, accuracy is best for random forest model , so Random Forest model is applied to predict the 20 quiz results

```{r}
predict_RF<- predict(model_randomForest, testing)
predict_RF
```
I have used assumption in processing data is to limit dataset for attributes that are non-zero in the Training and tEsting dataset. I haved faced lots of error about missing attributes in Testing data but available in Training dataset but this got tackeled by using this assumption.



